### 2022-06-24 class-info/template.qmd

---
title: Short Paper
author-list:
  - name: Alice Anonymous
    affiliation: Some Institute of Technology
    email: alice@example.com
  - name: Bob Security
    affiliation: Another University
    email: bob@example.com

abstract: |
  Duis autem vel eum iriure dolor in hendrerit in vulputate
  velit esse molestie consequat, vel illum dolore eu feugiat nulla
  facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
  praesent luptatum zzril delenit augue duis dolore te feugait nulla
  facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
  sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
  aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
  vel eum iriure dolor in hendrerit in vulputate velit esse molestie
  consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
  accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
  delenit augue duis dolore te feugait nulla facilisi.

category: Research
bib: template.bib
bibstyle: abbrv-doi.bst
csl: ieee-transactions-on-visualization-and-computer-graphics.csl
ieee-keywords: 
  - \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visu\-al\-iza\-tion techniques}{Treemaps}
  - \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization design and evaluation methods}{}
firstsection: Introduction

link-citations: false

format: 
  latex:
    template-partials: [partials/doc-class.tex, partials/title.tex, partials/before-body.tex, partials/after-body.tex]

pdf-engine: pdflatex

date: last-modified
---

This template is for papers of VGTC-sponsored conferences which are \emph{\textbf{not}} published in a special issue of TVCG.

# Using the Style Template

- If you receive compilation errors along the lines of `\texttt{Package ifpdf Error: Name clash, \textbackslash ifpdf is already defined}` then please add a new line `\texttt{\textbackslash let\textbackslash ifpdf\textbackslash relax}` right after the `\texttt{\textbackslash documentclass[journal]\{vgtc\}}` call. Note that your error is due to packages you use that define `\texttt{\textbackslash ifpdf}` which is obsolete (the result is that `\texttt{\textbackslash ifpdf}` is defined twice); these packages should be changed to use ifpdf package instead.
- The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the `\texttt{\textbackslash autoref\{reference\}}` call (instead of `\texttt{Figure\~{}\textbackslash ref\{reference\}}` or similar) since `\texttt{\textbackslash autoref\{reference\}}` turns the entire reference into an internal link, not just the number. Example: @fig-sample.
- The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use `\texttt{\textbackslash includegraphics\{CypressView\}}` (instead of `\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}`).
- For adding hyperlinks and DOIs to the list of references, you can use `\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}` (instead of `\texttt{\textbackslash bibliographystyle\{abbrv\}}`). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the `\texttt{http://dx.doi.org/}` url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (`\texttt{pdfendlink ended up in different nesting level ...}`) if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to `\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}` (at the expense of removing hyperlinks from the bibliography) and try `\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}` again after some more editing.\\[1em]
\textbf{Note 2:} the `\texttt{-narrow}` versions of the bibliography style use the font `PTSansNarrow-TLF` for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the [`paratype` package](https://www.ctan.org/pkg/paratype), and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a `\texttt{-narrow}` bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the `\texttt{-narrow}` in the file name.\\[1em]
DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.

# Bibliography Instructions

- Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
- Merge multiple references into one; e.\,g., use \cite{Max-1995-OMF,Kitware-2003} (not \cite{Kitware-2003}\cite{Max-1995-OMF}). Within each set of multiple references, the references should be sorted in ascending order. This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
- Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
- Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
- Articles in journal---items to include:
  - author names
	- title
	- journal name
	- year
	- volume
	- number
	- month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})

- use journal names in proper style: correct: "IEEE Transactions on Visualization and Computer Graphics", incorrect: "Visualization and Computer Graphics, IEEE Transactions on"
- Papers in proceedings---items to include:
  - author names
	- title
	- abbreviated proceedings name: e.\,g., "Proc.\textbackslash{} CONF\_ACRONYNM" without the year; example: "Proc.\textbackslash{} CHI", "Proc.\textbackslash{} 3DUI", "Proc.\textbackslash{} Eurographics", "Proc.\textbackslash{} EuroVis"
	- year
	- publisher
	- town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)

- article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
	- paper "Marching Cubes: A High Resolution 3D Surface Construction Algorithm"
	- should be entered as "\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm" or  "\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm"
	- will be typeset as "Marching Cubes: A high resolution 3D surface construction algorithm"

- for all entries
	- DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
	- provide full page ranges AA-{}-BB

- when citing references, do not use the reference as a sentence object; e.\,g., wrong: "In \cite{Lorensen-1987-MCA} the authors describe \dots", correct: "Lorensen and Cline \cite{Lorensen-1987-MCA} describe \dots"

# Example Section

Lorem\marginpar{\small You can use the margins for comments while editing the submission, but please remove the marginpar comments for submission.} ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam
voluptua [@Kitware-2003;Max-1995-OMF]. At vero eos et accusam et
justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est.

# Exposition

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat~\cite{Kindlmann-1999-SAG}.

\begin{equation}
\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

## Lorem ipsum

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

## Mezcal Head

Lorem ipsum dolor sit amet (see @fig-sample), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

### Duis Autem

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
ipsum dolor sit amet.

![A visualization of the 1990--2015 data. The image is from @Isenberg-2017-VMC and is in the public domain.](paper-count-w-2015-new){#fig-sample}

### Ejector Seat Reservation

Duis autem \cite{Lorensen-1987-MCA}^[The algorithm behind
Marching Cubes \cite{Lorensen-1987-MCA} had already been
described by Wyvill et al. \cite{Wyvill-1986-DSS} a year
earlier.] vel eum iriure dolor in hendrerit
in vulputate velit esse molestie consequat,^[Footnotes
appear at the bottom of the column.] vel illum dolore eu
feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
ut laoreet dolore magna aliquam erat volutpat.

#### Confirmed Ejector Seat Reservation

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat [@Nielson-1991-TAD]. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

#### Rejected Ejector Seat Reservation

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie

# Conclusion

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this-

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).