### 2022-06-24 partials/title.tex

$if(title)$
  \title{$title$}
$endif$
$if(subtitle)$
  \subtitle{$subtitle$}
$endif$

\author{%
$for(author-list)$%
  $it.name$\thanks{email: $it.email$}\\\scriptsize $it.dept$\\\scriptsize $it.affiliation$%
  $sep$\and%
$endfor$%
}

$if(date)$
  \date{$date$}
$endif$

$if(teaserimg)$
\teaser{
  \centering
  \includegraphics[width=\linewidth]{$teaser-img$}
  \caption{$teaser-caption$}
  \label{fig:teaser}
}
$endif$

$if(abstract)$
\abstract{$abstract$}
$endif$

\vgtccategory{$category$}

$if(ieee-keywords)$
\CCScatlist{
$for(ieee-keywords)$
$it$
$sep$;
$endfor$
}
$endif$