### 2025-03-17 index.qmd

---
title: A Guide to Designing Experiments to Test Statistical Graphics
author:
  - name: Emily Robinson
    dept: Statistics Department
    affiliation: 
    email: erobin17@calpoly.edu
    orcid: 0000-0001-9800-7304
    affiliations: 
      - name: Statistics Department, California Polytechnic State University
        address: Faculty Offices East, Building 25, Room 107D
        city: San Luis Obispo
        state: CA
        postal-code: 93407-0405
  - name: Heike Hofmann
    orcid: 0000-0001-6216-5183
    email: hhofmann4@unl.edu    
    affiliations: 
      - name: Statistics Department, University of Nebraska - Lincoln
        address: 340 Hardin Hall North Wing
        city: Lincoln
        state: NE
        postal-code: 68503
  - name: Susan Vanderplas
    email: svanderplas2@unl.edu
    orcid: 0000-0002-3803-0972
    affiliations: 
      - name: Statistics Department, University of Nebraska - Lincoln
        address: 340 Hardin Hall North Wing
        city: Lincoln
        state: NE
        postal-code: 68503

abstract: |
  In this paper, we discuss considerations and methods for experimentally testing visualizations. We discuss levels of user engagement with graphics, common issues when developing a sampling or data generation model, the importance of pilot testing, and data analysis methods. Along the way, we also provide recommendations of how to avoid some of the unique pitfalls of human testing in statistical and visualization research.

bibliography: refs.bib
csl: apa.csl

mainfont: Helvetica
monofont: Roboto
number-depth: 2
number-sections: true

execute: 
  echo: false

header-includes:
   - \usepackage[dvipsnames]{xcolor} % colors
   - \newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
   - \newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
   - \newcommand{\hh}[1]{{\textcolor{Green}{#1}}}

date: last-modified

format:  
  wordcount-pdf:
    keep-tex: true
    fontsize: 10pt
---

<!-- Importance of testing statistical graphics -->

# Introduction

<!-- \hh{How do you call the tested audience? viewer: 5  user: 32  mostly in the user engagement section, participant: 50+ times. -->
<!-- I'm not sure that I love the word 'user', but I can see how it is more active than participant. } -->

Data visualizations are a critically important tool for communicating scientific information to the public in what creators hope is an easy-to-digest, visually attractive form.
There are many strategies for creating charts and graphs, from Tufte-esque minimalism [@tufte] to charts designed with extra imagery and aesthetic appeal that draw the viewer's attention and persist in memory [@cairoFunctionalArtIntroduction2012]. 
For a specific type of data, there are also usually many different chart forms to display that data: for instance, if we have a set of categorical data and we wish to show the relative proportions of each category, we could do so using a stacked bar chart or the polar equivalent, a pie chart. 
There have been several attempts to list out all of the types of charts [@ribeccaSearchChartsData], create a taxonomy of charts [@bertin1983semiology;@desnoyersTaxonomyVisualsScience2011], and even to create charts using a domain-specific grammar of graphics [@wilkinsonGrammarGraphics1999] that is also useful for classification. 
One extremely useful reference is from Data to Viz (https://www.data-to-viz.com/), which uses a decision tree to show different visualizations compatible with the data; R, python, D3.js, and React code are provided to demonstrate how to create those visualizations. 
With all of the different design choices available, how are chart creators to know what is the best approach for communicating data to the appropriate audience?

While there are heuristics, rules-of-thumb, and guidelines [@asme-standards-graphics;@allenVisualizingScientificData2016;@haemerDoubleScalesAre1948;@fewInformationDashboardDesign2006;@kosslynGraphDesignEye2006] for creating useful and visually attractive data displays, the best way to establish the efficacy of various design decisions is to test the visualization on humans, evaluating different variants under controlled conditions [@clevelandShapeParameterTwoVariable1988;@clevelandGraphicalPerceptionGraphical1985].
Empirical assessments of visualizations, when carefully designed, allow statisticians to determine which representation of the same data is most effective along one or more dimension(s) of interest: estimation or prediction accuracy, within or between group comparisons, response time, and ability to draw real-world decisions are common goals for charts.

It is extremely challenging to design studies which strike the right balance between experimental control (i.e. internal validity) and generalizability to a wider context (i.e. external validity). 
Simply asking people to read quantities off of a graph may not generalize beyond the questions asked or the data used in the chart [@eellsRelativeMeritsCircles1926;@croxtonBarChartsCircle1927;@vonhuhnFurtherStudiesGraphic1927;@croxtonGraphicComparisonsBars1932], but designing a study that is sufficiently robust to those issues requires manipulation or control of so many factors that the amount of participants and trials quickly becomes daunting or unaffordable.
In addition, when conducting graphics experiments, researchers are in the unusual position of being both the subject matter expert and the statistician, providing an unusual amount of control over not just the experimental design but also the specific treatments, levels, and experimental protocols. 
The amount of choices required to develop, pilot, and run an experiment can be overwhelming. 
In this paper, we attempt to distill the experience gained from conducting several different types of graphics experiments [@clusters;@sineillusion;@emily-diss;@power;@vanderplasFramedReproducingRevisiting2019;@vanderplasEscapingFlatlandGraphics2024], discussing the use of different testing methods [@vanderplasTestingStatisticalCharts2020], the process of designing a graphical experiment, and analysis of the resulting empirical data. 
It is our hope that this paper will lower the barriers that exist for conducting empirical graphics research and reduce the probability of costly mistakes.

@sec-testing-methods discusses different methods for testing graphics, and which methods best address different levels of user engagement. 
In @sec-model-dev, we discuss the process of developing the data-generating model used to control the statistical features of data in the tested visualizations. 
Model development is a nuanced and iterative process that ultimately determines the success and generalizability of the experimental results. 
In @sec-exp-dev, we discuss the design of the experimental protocol - the choice of platform, number and type of trials, and flow of the experiment. 
We briefly consider different experimental design considerations in @sec-exp-design, but focus primarily on factors specific to graphics experiments, and then move to the importance of pilot testing in @sec-pilot-test. 
Finally, we  provide some common analysis strategies in @sec-analysis, including strategies for handling the unexpected data features which are so common in graphical testing experiments.

# Testing Methods and User Engagement {#sec-testing-methods}
<!-- Lineups, You Draw It, Direct Assessment -->

There are many different testing methods used to empirically assess statistical graphics. 
This paper uses studies conducted online  without additional equipment as primary examples, though many of the same considerations apply to in-person experiments conducted using additional equipment, including 3D printed charts, eye-tracking equipment, and interactive data displays.
Online experiments have lower overhead, offer relatively fast data collection, and provide useful results for well-designed experiments.
The toolkit used for these experiments is R-based [@R], and includes ggplot2 [@ggplot2] and Shiny [@shiny] as primary components.
In many experiments, we customized the Shiny interface with JavaScript and d3 [@d3], enabling interactive graphics, use of svgs, and other useful extensions.
While we prefer this set of tools, most of the observations described here apply to a wide variety of different workflows for graphical experimentation, including in-person experiments.

It is important to consider the level of user engagement which is necessary to complete a particular visual or graphical task. 
For instance, testing whether someone can detect an effect such as a linear trend in noisy data is a perceptual question. 
Perceptual questions are often examined experimentally using methods which allow the user to interact with the data on a basic visual level: users are presented with a visual stimulus and answer yes/no questions to indicate whether the effect is detected. 
Numerical estimation is another common task when testing graphics: in these experiments, the participant views a chart, estimates the requested numerical quantity, and enters the estimate into the application through a numerical input, slider, or other form element. 
Sometimes, it is possible to set up a scenario where the user adjusts the plot using a set of controls designed to provide a fixed set of interactive operations. 
This type of user engagement was used to assess the strength of the sine illusion [@sineillusion]: users adjusted the strength of a transformation designed to correct the illusion until the lines appeared to be the same length, as shown in @fig-sine-illusion, providing a direct measure of the magnitude of the sine illusion's effect. 
In other situations, it may be preferable to have the user directly interact with the visual stimulus. 
In @vanderplasEscapingFlatlandGraphics2024, participants were asked to rotate and interact with a 3D rendered bar chart; the application recorded user interactions and corresponding rotation matrices, providing insight into the visual comparisons the user may have been performing. 
This information was used as a supplement to the explicitly provided estimates, providing some contextual information as well as the ability to identify the level of participant engagement with the questions.
When experiments are conducted as part of classroom experiential learning, it is sometimes helpful to be able to separate the low-effort participants from those who were fully intellectually engaged in the task. 
Interactive graphics provide another level of user engagement that can be much more open-ended. 
With interactive graphics, researchers can ask participants to directly annotate plots, toggle aesthetics, and highlight groups and plot features.
Careful implementation of the experiment application may allow for each of these interactions to be recorded and analyzed, producing a rich, if messy, set of data that may allow researchers to tease apart visual estimation error from common shortcuts such as rounding used during direct numerical estimation.

![Direct adjustment of a plot in a perceptual task. In this experiment, designed to assess the strength of the sine illusion, the user adjusts the plot using - and + buttons, which control the strength of a transformation designed to correct the effect of the sine illusion. When the user is satisfied that the lines are of equal length, they select the 'Finished' button to move to the next task. The experiment used a psychophysics experimental design, the method of adjustment, but leveraged the interactive Shiny interface to record the entire sequence of adjustments made by the user for each trial. A demo version of this application can be found at https://shiny.srvanderplas.com/sine-illusion/.  ](sine_illusion_screenshot.png){#fig-sine-illusion}

Visual inference [@bujaStatisticalInferenceExploratory2009;@wickhamGraphicalInferenceInfovis2010] is another useful testing tool for perceptual questions such as "which chart displays this data more clearly" [@power] while simultaneously assessing the statistical significance of the graphical finding in a chart. 
Visual inference charts are often called "lineups" in analogy to the criminal procedure where the suspect is placed in a line with several other individuals with similar characteristics. 
In a graphical lineup procedure, there is a target plot containing the real data, embedded in an array of innocent "null" plots generated through resampling or simulation. 
If viewers consistently pick the target plot at a higher rate than any of the null plots, the target plot is said to be visually significant [@loyDiagnosticToolsHierarchical2013;@majumderValidationVisualStatistical2013] and a "see" value, the visual analogue of a $p$-value [@niladriroychowdhurySeeValueApp2020], can be calculated using the `vinference` R package or the process described in @vanderplasStatisticalSignificanceCalculations2021. 
The details of this calculation are beyond the scope of this broader discussion of how to test charts, but more detail on visual inference is provided in \<insert citation to visual inference WIRE article under development\>. 
<!-- Beyond the quantitative assessment of the signal with $p$- and see-values, the human involvement also allows us to collect information on the qualitative nature of *what* a viewer sees.-->

In another variation of the statistical lineup procedure, data generated from two models are compared, with target plots from each model embedded in the array of $K$ total plots. 
The $K-2$ null plots are constructed from a mixture model that blends the two competing models [@clusters]. 
Viewers are asked to select the panel(s) which are most different, and the primary source of information are trials in which viewers selected the target from one model but not the other, indicating that the display method used allowed viewers to differentiate one model's data (but not the other) from the nulls created through a mixture model. 
This variation allows the experimenter to assess graphical design choices to determine whether they effectively emphasize structural differences in the data [@clusters].

One advantage of the visual inference technique is that the experimenter can ask a very general question, such as "which of these plots is the most different?", rather than a specific question about the displayed data which may require more quantitative sophistication. 
All of the necessary information to make the decision is embedded in the choice of the model used to generate the null plots. 
This feature is extremely convenient when conducting the experiment and even allows small children to complete the task. 
The downside is that as a result, visual  inference experiments do not allow experimenters to assess the viewer's understanding of the information shown in the chart. 
In most cases, visual inference experiments remove any contextual information from the charts, including axis labels and values, plot titles, and so on, in order to encourage participants to make decisions based solely on the graphical presentation. 
This lack of context is a double-edged sword: visual inference can involve participants who do not have any mathematical training or instincts (including children), but researchers also cannot use this technique to assess higher levels of engagement with a chart, such as estimation, prediction, or reasoning based on displayed information.  
<!-- An example of a statistical lineup for testing the perception of exponential growth is shown in @fig-lineup-log. -->

<!-- ![A lineup experiment testing whether users can perceive differences in rates of exponential growth on linear and log scales. In addition to asking users to select the plot which is the most different, we also asked for basic reasoning and the user's level of confidence in their answer.](Lineup_log_screenshot.png){#fig-lineup-log} -->

To assess the viewer's *understanding* of information shown in a chart, we must ask questions and allow the user to provide feedback. 
User feedback may be collected on a numerical scale or through the use of written comments, recorded "think-aloud" processes, and other more qualitative interaction methods. 
In some studies, asking users to interpret a chart within a larger scenario can be effective, as in @fig-estimation-describe, while in others it is more helpful to ask users to explain answers. 
In visual inference studies, asking users why a specific panel was chosen has been demonstrated to provide rich insight into otherwise confusing numerical results [@clusters].

Think-aloud methods ask the viewer to narrate their internal thought process, either during or after completing a task [@thinkaloud]. 
These recordings (or transcripts) can provide valuable insights into conscious cognition, and are often used when conducting usability studies. 
While we have not to date recorded users talking out loud about what they are seeing during a study, think aloud methods could easily be implemented within a Shiny application, with audio recordings saved to the server for transcription and analysis
[@dunbar1995scientists; @traftonTurningPicturesNumbers2000; @kirschenbaum2003comparative].
It is even possible that these recordings could be automatically transcribed using speech-to-text models.
We have used think-aloud methods informally during pilot studies to "harden" graphical experiments and verify the selection of parameters used in an experiment. 
The success of this approach, combined with the few studies which used think-aloud to assess charts [@kulhavyCartographicExperienceThinking1992;@leeHowPeopleMake2016;@haiderStrategiesDetectingDifference2021], suggests that think-aloud methods are an often-overlooked but useful tool for assessing data visualizations.

![This question asks users to write out a description of how the population of Ewoks changes over time, without any further cues, to determine whether participants default to multiplicative or additive language descriptions.](Estimation_describe_plot_crop.png){#fig-estimation-describe}

Of course, in an online, asynchronous experiment, every user interaction with the testing materials (typically hosted on a web page) can also be recorded along with time stamps, mouse positions, browser size and screen resolution, and other information. 
While we have not used this type of information heavily in our experimental analyses thus far, in most experiments we collect time stamp data in order to assess how long participants spend on each question. 
Typically, the first round of test questions takes the longest for participants to complete. 
Additional replicates do not usually affect accuracy (i.e. there is no immediate learning effect) until after `too many' tests cognitive fatigue proves to be detrimental to accuracy [@chowdhury2018]. 
This sweet spot between replicates and fatigue depends on the cognitive burden in each test and should factor into designing the experiment.
<!--While we have not used this type of information heavily in our experimental analyses thus far, we do collect some time stamp data in order to assess how long participants spend on each question; it could almost certainly be more informative in the right setting.-->
In some experiments, we have provided participants with supportive tools, such as "scratch pads" and calculators built into the Shiny application to support the complex calculations required to answer higher-level numerical estimation questions (@fig-estimation-calc). 
In order to be supportive, the tools must be easy to use, but assuming this bar is met, the tools can reduce participant cognitive load while recording a wealth of information.
This information provides real insight into how participants were looking at the data, what strategies they tried and discarded for reading the chart, and what visual estimation methods were used. 
While systematic analysis and modeling of this data may be difficult, as it is usually messy and often must be manually coded, the insights provided can be extremely useful. 
However, unless participants are required to use these tools, it is difficult to gather comprehensive information - those participants that don't use supportive tools likely differ in meaningful ways from those who do.
As a result, the information gathered from supportive tools likely does not generalize to the entire sample.

![This question asks participants for a numerical estimate, but provides a basic calculator and scratchpad. All user interactions with the calculator and scratchpad are logged, providing insight into the user's thought process and estimation strategy.](Estimation_numerical_screenshot_crop.png){#fig-estimation-calc}

One of the most difficult components of designing an experiment which asks users to directly estimate information from a chart using a full scenario (background information, etc. as well as contextual details from the chart) is that the questions must be extremely carefully constructed.
Mathematics education researchers provide guidelines for selecting different levels of questioning in order to assess graph comprehension: literal reading of the data, reading between the data, and reading beyond the data [@wood1968objectives;@curcio1987comprehension;@friel2001making;@glazer2011challenges]. 
In a recent study, we identified questions based on this framework to evaluate direct estimates and extend those estimates to make comparisons between two points.

Even when great care is taken with the construction of the question, participant answer accuracy is fundamentally limited by the fact that many participants do not read and interpret the question with the care and precision that it was written.
Questions that ask participants to e.g. estimate the multiplicative change in a quantity at two time points may be misunderstood as asking for an estimate of the additive difference, and the resulting estimates are then one or more orders of magnitude off of the correct answer.
This is one area where lineup methods are convenient - they do not depend on participants to understand the nuances of language or scenarios built around the chart under investigation.
However, in some situations it may be sufficient to ask participants to estimate direct numerical quantities that have little contextual information, as done in [@vanderplasFramedReproducingRevisiting2019] when assessing the accuracy of framed plots re-created from the Statistical Atlas.

Another useful measurement strategy is to require participants to *engage directly* with an interactive visualization. 
This is useful in a directed task, where users are asked to interact with the chart in a specific way and the result is recorded, but it is also possible to use interactive visualizations in an open-ended task, recording how users engage with the graphic in an exploratory (as opposed to goal-directed) manner. 
In one recent experiment, we asked participants to forecast an exponential trend, with data presented on either a linear or log scale. 
Using JavaScript code modified from New York Times interactive graphics "You Draw It" features [@katzYouDrawIt2017], we had users draw trend lines with their computer mouse and make forecasts directly on interactive charts, with the data and user-drawn predictions recorded to our database.
With interactive graphics rendered using JavaScript (or other web libraries), the only limit to the types of questions one can ask in testing graphics is one's ability to write code to interact with the visualization library. 
This type of testing method can be extremely natural for participants, but it also is hard to generalize when discussing testing methods because of the potential range of applications where it might be employed.

Whichever testing method is chosen should be appropriate to the type of question under investigation and the level of visual and cognitive engagement required to answer that question. 
While lineups are excellent tools for assessing perceptual questions, they cannot address questions aimed at understanding how people use charts within the wider context of a story or practical task; this requires more direct methods with higher ecological validity.

<!-- \ear{kind of a sudden jump into this paragraph, but I'm not sure how to transition into it right now.} -->
All of the testing methods described here require significant work to develop a strategy for data generation appropriate for testing the underlying question. 
For instance, when testing the perception of exponential growth, we had to develop a model which would generate data with varying growth rates, but where the data had a pre-specified domain and range. 
The data generating model is particularly critical when using lineups, as the null sampling model must replicate the important visual features in the data. 
Each testing method has specific requirements, but it is important to carefully calibrate the model parameters to allow for some variability, but not too much, and to ensure that participants can succeed at the task and do not feel like they are being made to analyze random noise. 
This Goldilocks-style problem is the focus of the next section.

# Experiment Development Life Cycle

Developing a graphics experiment is often a highly iterative process, but it can help to approach the design process by first optimizing the model and data generation method before spending time on optimizing the specific stimuli or customizing the data collection platform.
This is important because the model parameters and data generation process inform the experiment structure and thus impact decisions made downstream.

Once the model and data generating mechanism are set, it is useful to revisit the primary questions of interest and determine how to measure the responses effectively. 
Secondary measures, such as response time, free responses, and confidence level should also be determined. 
These choices will inform the choice of a data collection platform and may also inform the participant recruitment method.

Next, we recommend developing a preliminary data analysis plan, specifying the general category of model which will be used (e.g. generalized linear mixed-effects model, t-test) and the contrasts which are most interesting. 
This sets up the experimental design decisions, but also ensures that as the data collection platform and process is developed, any design constraints are considered. 
Development of the data collection application is the next step, using draft graphics and the set of participant response measures of interest.

There are at least 3 stages of testing in a graphics experiment: informal tests, a pilot study, and the main experiment. 
The informal tests are critical for identifying issues with the data collection application, but can also be used to calibrate the number of tasks required of each participant. 
As the number and complexity of tasks increases, the number of trials we can ask participants to complete during a session decreases. 
The informal testing stage allows researchers to consider the tradeoffs inherent in the decision to reduce the amount of information collected for each task, reduce the number of tasks, or mitigate participant fatigue in other ways.

During preliminary testing, we use an optimistic number of trials per participant, so that we can determine when participants become overly fatigued. 
For instance, we might ask test participants to evaluate 20 graphical lineups (400 total plots), even though we expect to reduce the number to 10 or 15 during the main experiment. 
We test the application in individual or focus group sessions, often using graduate students, colleagues, social media acquaintances, and conscripted family members. 
After these participants complete the study, we ask questions about fatigue to determine what range of trials per participant is reasonable. 
At the end of our initial experiment tests, we have enough information to determine the basic parameters of the experimental design (e.g. how many blocks in an incomplete block design can we have with the factors under investigation). 
The number of trials a single participant can complete without excessive fatigue impacts the number of blocks and the strategy by which we allocate trials to each participant.

In addition, we must consider how long participants take to complete the required number of trials.
Completion time is used to determine participant compensation (if using a participant recruiting platform). 
Ethics boards and some recruitment platforms require that participants are paid a reasonable wage for their time (currently, around \$15 US per hour), and platforms may ask for median completion time and automatically reject submissions from participants who are too far under or over the specified time limits; they may also require additional participant payments if the median time estimate is too far below the actual average completion time during the experiment.
Platforms may also calculate fees based on both the participant payment and number of participants recruited, with additional fees to recruit e.g. demographically representative samples; as a result, it can be advantageous to balance cognitive load concerns with the fee structure used by the selected recruitment platform.

The findings from the initial test of the experimental procedure are then used to revise the data collection procedure in preparation for one or more pilot studies.
It is important to ensure that the software platform, trial allocation, and other components of the experiment are functioning as desired before a formal pilot study is conducted. 
In some cases, the pilot study is as simple as a "soft launch" of the main experiment, where the total number of trials is pre-specified and only a few trials are released initially to ensure that data collection works as expected. 
In others, the pilot study is conducted first, and results from that study are used to determine the sample size for the main experiment. 
At this point, data collection, analysis, and reporting proceed much as in any other experiment.

# Developing a Model {#sec-model-dev}

Once the graphical task has been identified, it is necessary to develop a model which can be used to explore the graphical features of interest in a precise manner. 
This is the single longest part of the entire experimental design and execution process, in part because choosing a model that replicates important visual features of the data is  extremely complex [@hullmanDesigningInteractiveExploratory2021;@cookFoundationAvailableThinking2021;@vanderplasDesigningGraphicsRequires2021].

There are two main options when developing a statistical model for graphical testing: start with a large data set and sample from that data set [@power], or start from a model and sample data from that model generating process [@sineillusion;@clusters;@emily-diss].
This decision is largely determined by the availability of a large data set containing the requisite features of interest and the qualities being manipulated in the experiment. 
For instance, @power used samples of different sizes from a pre-existing data set to manipulate the amount of signal in each comparison; with a small sample, there is less signal and the same amount of noise, making the true plot harder to spot. 
In many situations, though, a convenient data set with the right properties is harder to acquire, and it becomes necessary to develop a sampling model to generate data for user evaluation.

The tools we discuss in the remainder of this section can be applied both to pre-existing data sets and to model-based sampling methods.

## Screening Parameters with Simulation

The choice of the parameter space used in testing is crucial to gain insight from a study without putting too much burden on participants with overlong studies. 
Choosing an appropriate space for testing parameters is a well-known problem in psychometric testing: the space considered should cover the area between 'only some activation' to 'almost full activation' of an appropriate psychometric function [@schuttPainfreeAccurateBayesian2016;@valentinDesigningOptimalBehavioral2024]. 
When testing charts, visual assessment is obviously key, but researchers can make use of statistical indices related to the testing condition to narrow the parameter space to a reasonable and efficient subset from which maximal information can be acquired.  
These statistical indices may also serve as quantitative proxies for the difficulty of the visual task. 
To identify a statistical proxy for visual difficulty that may help with narrowing the parameter space, it can be useful to consider numerical measures used to estimate the same types of visual information that will be assessed in the experiment. 
For instance, we have used

- $R^2$ as a measure of the strength of a linear relationship
- Gini inequality as a measure of the strength of clustering
- lack-of-fit statistics to assess the amount of curvature in an exponential relationship (shown in @fig-lof-density-curves)

Then, a wide range of potential combinations of parameter values or sampling strategies can be explored and summarized graphically; if the numerical statistic cannot differentiate between the null and target under a condition, it is reasonable to think that a visual inspection of the data may also not show significant results.
As with any measure, it is important that difficulty levels span a range from easy to hard; we do not learn anything from finding out that everyone can distinguish all of the combinations.
This portion of the design is somewhat analogous to selecting a range of doses of a chemical in a dose-response experiment.

```{r lof-density-curves, echo = F, include = T, message = F, warning = F}
#| label: fig-lof-density-curves
#| fig-cap: "Density plot of the lack of fit statistic showing separation of selected difficulty levels: High (obvious curvature), Medium (noticeable curvature), and Low (almost linear). Each density plot is the result of 1000 simulations from a model $y_i = \\alpha\\cdot e^{\\beta\\cdot x_i + \\epsilon_i} + \\theta$, where $\\epsilon \\sim N(0, \\sigma^2)$. $\\alpha$ and $\\theta$ were selected after manipulation of $\\beta$ and $\\sigma$ to ensure that all data generated had similar $y$ ranges so as not to provide visual cues about model differences outside of the plot curvature."
#| out.width: "\\columnwidth"
#| fig.height: 2.5
#| fig.width: 5

library(tidyverse)
lofData <- read.csv(file = "https://raw.githubusercontent.com/earobinson95/EmilyARobinson-UNL-dissertation/main/data/01-lineups/lineup-lof-data.csv")
lofPlot_curvature <- lofData %>%
  mutate(Curvature = factor(Curvature, levels = c("Obvious Curvature", "Noticeable Curvature", "Almost Linear"), labels = c("High Curvature", "Medium Curvature", "Low Curvature"))) %>%
  mutate(Variability = factor(Variability, levels = c("Low"))) %>%
  ggplot(aes(x = statistic, fill = Curvature, color = Curvature)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual("Difficulty", values = c("#004400", "#116611", "#55aa55")) +
  scale_color_manual("Difficulty", values = c("#004400", "#116611", "#55aa55")) +
  theme_bw(base_size = 14) +
  theme(legend.position = "bottom",
        axis.text    = element_text(size = 10),
        axis.title   = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 10),
        legend.key.size = unit(0.5, "line")
        ) + 
  scale_x_continuous("Lack of Fit Statistic") +
  scale_y_continuous("Density") + 
  theme(legend.position = c(1, 1), legend.justification = c(1.01, 1.01), legend.box.background = element_rect(fill = "white", color = "black"))
lofPlot_curvature
```

While this method is certainly more critical for model-based sampling methods, it is also important when data are generated by sampling from a larger data set. 
When sampling from a larger dataset, parameters are more often sample size and stratification methods, but it is still important to iteratively assess the data generating procedure through simulation. 
Using numerical proxies for visual characteristics of data displays such as curvature, linearity, scatter, dispersion can assist with identifying optimal parameter settings to use across different experimental conditions. 
Even with this strategy, it is still critical to fine-tune the parameter choices with visual calibration and pilot testing.

## Fine-Tuning Parameter Choices

Once an appropriate set of parameters are identified using the numerical screening method,  it is important to calibrate these parameter selections visually. 
No numerical statistic is a perfect measure of what we actually see: at best, they are approximations of what we might potentially see. 
We have found it to be useful to have one experimenter calibrate the model parameters at a gross level, and then have another experimenter narrow in on the parameters which are visually reasonable within the selected range. 
Then, both examiners visually inspect a large number of plots generated using those parameters to get a sense for how difficult the task at hand is (this strategy is also described by @luModelingJustNoticeable2022). 
At some point, all experimenters become so visually saturated with the nuances of the data generating mechanism that it may become necessary to "sanity check" the protocol with family members, friends, and colleagues. 
These informal focus groups  provide extremely useful feedback and can help to counteract the visual saturation of being immersed in the design of a visualization experiment for months at a time.

## Visual Assessment is Critical

We cannot overstate the importance of visual assessment of your model stimuli, preferably with fresh eyes. 
We highly recommend performing several rounds of think-aloud pilot testing (focus) before deploying an experiment. 
In support of this assessment, we offer up a cautionary tale of our own experience: that of @clusters, where we designed an experiment to test which plot aesthetics promoted discovery of linear trends and/or clusters.

The experiment was a 2x3x3 factorial exploration of three data generating parameters, with 3 replicates at each parameter combination (54 data sets) and 10 aesthetic combinations (for a total of 540 lineups). 
Each lineup had 20 different sub-panels, so we should have carefully visually inspected some 10,800 different panels. 
As is evident from the fact that we're telling this story as a cautionary tale, we missed a critical problem with our data-generating mechanism: when clusters were assigned to randomly generated data after the fact, we didn't control the cluster size, leading to clusters of one or two points in relatively few sub-panels. 
This became particularly noticeable when bounding ellipses were added to the plot, as the method used to generate those ellipses required at least 3 points in the cluster. 
The missing boundary ellipse in the corresponding sub-panels escaped our notice during the stimuli proof-reading phase of the experiment, but did not escape the notice of our participants, who only needed to examine about 10 lineups each (around 200 panels). 
An example of one of the problematic lineups is shown in @fig-lineup-problems: many participants selected panel 16 because of the missing ellipse; not a wrong choice, but certainly not the effect we intended to test.

![A lineup from Vanderplas & Hofmann [-@clusters]. Panel 10 shows the clustered target data and panel 17 shows the target data with a strong linear relationship; either of these target panels was the expected choice. Unfortunately, panel 16 has only two bounding ellipses shown, which is an unintentional difference that resulted from a faulty method for assigning clusters to null plots; many participants selected this panel instead of one of the target panels.](lineup-missing-ellipse.png){#fig-lineup-problems}

One reason why it is so difficult to generate sampling models for visual explorations is that our visual system is optimized for identifying differences between groups.
This ability can interfere with the natural  to use the null sampling models that might be used in equivalent numerical tests when running experiments that use visualizations. 
We re-ran the experiment using a different clustering method that controlled the number of points in each group. 
Instead of noticing the number of ellipses, participants instead used the differences in size and shape of the ellipses formed when clustering after the data generating procedure. 
That is, participants could still detect the artificial nature of the induced clusters using other features. 
While it can be difficult to get the data generating method right, it is essential to conducting visual experiments that generalize well beyond the effects shown in a single data set or phenomenon. 
The time and effort invested in this step at the outset of the experiment pays dividends when it allows for clear generalization of the experimental results to an entire statistical concept rather than a single data set.

# Protocol Development {#sec-exp-dev}

It would be difficult to develop a full data generating model without some idea of the experimental protocol: the basic equipment required for the experiment, some idea of what questions users will answer, where and how data will be collecteed, and so on. 
These experimental design factors are fairly natural for scientists to accumulate over the course of imagining and planning an experiment. 
When conducting graphical tests, however, there are additional considerations beyond those taught in a standard experimental design course.
Experimenters must carefully consider how much information participants should have about the experiment, what platform to use to recruit participants, and the experimental infrastructure underlying data collection. 
In addition to our brief overview of considerations in our research, @kochariConductingWebBasedExperiments2019 has many helpful suggestions for conducting web-based cognitive and perceptual studies that also apply to statistical graphics experiments.

Of course, more standard statistical design considerations, like blocking, randomization, sample size, and analysis methodology are also important; we will discuss these briefly in @sec-exp-design. 
Here, we focus primarily on the experiment development process, with critical decisions in a roughly temporal order.

## Infrastructure

We have conducted visualization experiments using a wide variety of tools: custom web servers running interactive, PHP-based forms, generic web survey platforms (Qualtrics, Google Forms) for static graphics, and Shiny applications that control every part of the experiment interface (instructions, generating completion codes for participant payment, rendering interactive graphics, and generating fully randomized data for each participant). 
In our research, Shiny has provided the right balance between control over the experimental setting, procedure, etc. and the intricate details of web server administration and management, however, this balance is likely different for every lab and potentially for every experiment.

When using Shiny to collect data, we store participant response data and experiment parameters using SQLite tables, which are automatically synchronized with cloud storage and tracked with version control. 
This ensures that we have incremental records of the tables during data collection, and that all data is stored across multiple locations, guarding against hardware failures. 
In some studies, data is unique to each participant; in these cases, we highly recommend saving all generated data to a database as well, so that it is possible to go back and examine responses alongside the data used to generate the graphical stimuli.
Hard drive space is extremely cheap relative to almost any other cost in an experiment; saving all of the data is a sensible measure.

## Participant Recruitment

There are several different, commonly used methods for recruiting participants for visualization experiments and cognitive experiments more broadly. The selection of participant recruitment method depends on the infrastructure which will be used in the experiment, but the choice of participant pool may be more critical to the experiment results than the infrastructure and modality [@uittenhoveLabTestingWebTestingCognitive2023].

We have used each of the following strategies in experiments:

- Validated, representative panels of participants offered by specialized polling groups. In the US, this includes [NORC's AmeriSpeak](https://amerispeak.norc.org/us/en/amerispeak/about-amerispeak/panel-design.html) and [Gallup Panels](https://www.gallup.com/analytics/318911/us-social-research.aspx)

- Recruitment platforms designed for crowdsourcing research tasks, including  [Amazon Mechanical Turk](https://www.mturk.com/) and [Prolific](https://www.prolific.com/). Researchers may be able to obtain representative samples measured against census data along variables such as race, age, and sex, but participants are often more technologically sophisticated and educated than the general population on these platforms.

- In-person studies. In person participants are commonly recruited from undergraduate students, but in some experiments it may be preferable to recruit participants from the local community. Students are often convenient for academic studies, as participation in experiments is often a component of introductory course experiential learning activities, as in @vanderplasEscapingFlatlandGraphics2024. While the data generated from required classroom experiments may have higher variability, and it may be hard to generalize findings beyond undergraduate students, these studies can be conducted much more cheaply than studies conducted through online platforms or panels. Recruiting participants from the community for in-person research is also viable, but can be much more complicated, however, some topics, including research involving persons with specific disabilities, the elderly, or subject-matter experts may require recruiting participants outside the university.

- Social media and email-based recruitment. Researchers may post directly to social media sites such as Twitter/X, Mastodon, and BlueSky, forums and discussion sites such as Reddit, or general email lists that may be purchased from universities and other marketing organizations. If the goal is to recruit a specific population, such as meterologists or forensic examiners, forums and email lists may be an extremely effective way to recruit participants. We have used the [SampleSize subreddit](https://web.archive.org/web/20250228110244/https://www.reddit.com/r/SampleSize/) successfully and obtained participants that, while younger and more highly educated than those recruited from a platform, had similar results on visual estimation tasks [@vanderplasFramedReproducingRevisiting2019]. When using samples obtained from acquaintances, researchers should try to obtain more diversity than was present in historical papers, such as  @clevelandGraphicalPerceptionTheory1984, where the authors sampled their colleagues (and their wives).

@riceTestingPerceptualAccuracy2024 recruited participants using nationally representative panel samples and found that conclusions from fully representative samples of the population can be very different from volunteer samples recruited using other services, as many people are unmotivated to engage with charts, don't know how to read charts, or impose pre-formed conclusions onto visual displays. 
While using representative panel sampling products offered by organizations such as NORC or Gallup is more expensive than samples from Amazon MTurk or Prolific, these results may imply that results from less representative participant recruitment methods may not generalize to the broader population.
People who participate in studies via online platforms are more technologically sophisticated and educated than the general population, and this bias may significantly impact the conclusions. 
However, it is also reasonable to argue that higher education might make someone more likely to use charts and data for decision-making.  
Researchers designing a study should carefully consider whether the goal of the study is to generalize results to the adult population or to a subset of that population who make decisions based on data.

If data will be collected online, then we recommend considering the pros and cons of panel-based sampling methods and  online recruitment platforms. 
While Amazon MTurk was once the only large platform for this type of research [@heerCrowdsourcingGraphicalPerception2010], many researchers now prefer Prolific because its participants appear to be more attentive to tasks [@albertComparingAttentionalDisengagement2023;@peerDataQualityPlatforms2022]. 
Prolific's focus on academic research rather than developing training data for machine learning and AI means that its policies are tailored for these types of projects and its users may be more interested in science than those on other platforms. 
Experimenters should consider available options, compare pricing structures (as these vary widely), and consider whether add-on fees for e.g. demographically representative samples are worth the additional cost. 
It is also important to ensure that the platform supports the type of user engagement required for the study. 
Online recruitment platforms are more flexible than many panel survey platforms and allow experimenters to use a much wider range of stimuli and experimental designs, but this sacrifices some ability to generalize results to a wider population.

## Participant Instructions

It can be extremely helpful to include "practice" demonstrations of the task to show the basic process, logic, and reasoning. 
While it is tempting to make these tasks fully representative of the type of judgement which will be required of participants, practice tasks which are too close to the experimental task may bias participants.
We have found that it works well to have a relatively easy practice task which utilizes a slightly different type of plot and/or type of data than what will be tested in the experiment. 
In cases which require interactivity, gif animations of the task being carried out are useful, as are additional visual cues, such as the yellow box used in the `You Draw It' task[^1] to indicate that there were points which were not completed. 
Demonstrations can reduce cognitive load, but it is often critical not to prime participants to focus on the specific effect manipulated during the experiment. 
Finding the right examples and instructions to use in an experiment is a delicate process: priming participants can reduce the generalizability or relevance of the experiment, but if participants are confused about what question to answer or how to complete the task, response variability and participant cognitive load may be too high to detect an effect.
Pilot testing with a small group of volunteers who think aloud while completing the experiment can be extremely helpful in identifying problems with participant instructions. 
However, it is important that some pilot participants have lower levels of mathematical and statistical training that are similar to the participant population.
One technique that we have used is to wrap pilot testing into a presentation about the experiment to a general audience, perhaps as part of an undergraduate course or undergraduate research experience recruitment activity. 
The experiential component increases participant understanding of the research before the presenter explains the scientific goals behind the project, and it is often easy to get feedback on the experiment at the same time.

[^1]: See a gif of testing with `You Draw It' [here](https://i.imgur.com/GM5YSen.gif)

## Attention Checks and Answer Validation

In a perfect world, all participants would be fully engaged in the experiment, focusing only on that task from start to finish without interruption or distraction. 
In this world, participants would attentively read the unambiguous instructions for completing the experiment and executing these instructions flawlessly. 
Unfortunately, we do not live in this perfect world. 
Here, we focus primarily on mechanisms which can be used to exclude data based on participant noncompliance, rather than mechanisms which are used to automatically withhold participant compensation.  
There are several mechanisms that can be used to guard against or at least identify participants who are not fully engaged in the experiment or have severely misunderstood the instructions:

- Attention checks - questions inserted into the experiment which appear similarly to actual trials but which instruct participants to select a specific response or complete a trivial task (e.g. select the result of 2+2). Participants who do not successfully complete a certain percentage of attention check tasks may be excluded from the experiment. More guidance on attention checks can be found in @muszynskiAttentionChecksHow2023.

- User input validation - mechanisms which reduce the potential parameter space of user inputs to a valid or reasonable set of parameters. Examples include not allowing participants to select negative variance values, requiring continuous user-drawn lines on an interactive chart, and even checking to ensure that the user's input is of the correct type. Validation mechanisms may also ensure that participants answer all questions before proceeding to the next page.

- Time-based checks - @kochariConductingWebBasedExperiments2019 suggests that in tasks with longer instructions, it may be useful to remove participants who did not spend a certain amount of time on the instructions page. Similarly, it may be justifiable to remove participants who completed the experiment in an unreasonably short or long amount of time. On the other end of the spectrum, participants who take several hours to complete an experiment with a median completion time of 15 minutes may have experienced technical issues or been distracted.

Pre-specifying conditions where participant data will be removed from the study before analysis is critical for predictable issues, in order to streamline the analysis and defend against p-hacking concerns. 
When participants provide anomalous responses that are identified after the experiment, it is much harder, but not impossible, to justify removing the data. 
Justifications might include that no reasonable viewer would have produced a certain response, as in @sineillusion, or that the response most likely occurred due to a data entry error or a fundamental misunderstanding of the question, but this is only viable if the data is an egregious outlier.
In less severe cases, it is often preferable to leave the data in, increasing estimate variability but avoiding the need to justify and defend data cleaning decisions.

## Demographic Data

It may be useful to ask a few more demographic questions about STEM education level for studies which ask more of participants from a mathematical standpoint; while lineup studies have not found strong associations with those variables, lineup studies also do not require participants to engage with the data presented in a chart in a way that requires higher-order mathematical reasoning.
This has allowed us to make the argument to the ethics committee (institutional review board, or IRB) that our research is exempt, as we do not collect enough demographic information to identify participants, however, collecting reduced demographic information occasionally comes at a cost. 
One recent study examining the use of log scales and exponential data was conducted using Prolific, which recruits participants from around the world; we required only that participants were fluent in English to participate. 
It was only after the experiment was completed that we realized that different countries introduce logarithms as a concept at different points during primary and secondary education; it might be that individuals in some countries have much more experience with log scales than those educated in the United States. 
We mention this only to point out that while every experiment contains a few missed opportunities, it is worth giving careful thought to the demographic questions asked of participants and what information may be helpful during the analysis stage.

## Ethics Review
Once the protocol is developed, researchers generally have enough information to get approval from the ethics board to conduct an experiment on human participants. 
We are most familiar with regulations in the United States and can speak to that general process. 
Most graphics experiments conducted in the US fall under the "exempt" category of experiments which require only basic review and approval. 
These experiments record demographic information at a level that is not individually identifiable, ask participants to complete tasks that involve no risk, and experimenters record information which would not be embarrassing to participants if exposed. 
In order to ensure that our experiments fall into this category, we will often ensure that demographic information and participant responses are collected and stored separately from any identifiable information, such as e.g. user IDs from the participant recruitment platform that allow us to monitor task completion and pay participants for their time.
Researchers should also ensure that they comply with privacy laws such as the European [General Data Protection Regulation (GDPR)](https://gdpr-info.eu/chapter-3/); complying with this law while collecting data which is not identifiable can be complicated. 
We recommend consulting with your ethics board and institutional recommendations in order to maintain legal compliance and safeguard participant privacy.

# Experimental Design {#sec-exp-design}

In statistical design terms, most of our studies involve some type of balanced incomplete block design, where participants are assigned to a subset of experimental conditions which allow for estimation of the full range of effects specified in the model. 
The particular structure of these designs depends heavily on the factorial structure of the study and the contrasts of interest. 
In some experiments, it is important that participants see each set of data only once, while in others, it is critical that participants see the same set of data represented graphically under multiple conditions, at which point it becomes necessary to manipulate the order of trials to ensure that participants do not have back-to-back trials using the same data, because that might influence the responses. 
When randomization and blocking are used in the experiment, it is often beneficial to simulate the experiment before it is conducted to ensure that the data collection platform is correctly balancing and randomizing trials. 
Of particular concern is that in online experiments with demographic controls, some categories fill faster than others. 
In extreme circumstances, this can result in imbalances in block allocation within subgroups, which may complicate analysis of demographic data. 
It is preferable to iron out these issues before the main study begins, rather than trying to patch the data collection software in the middle of the study.

It is difficult to offer specific advice on the number of participants to include in the study, because the experimental design, participant engagement level, and the specific factors under investigation have such a large effect on the number of trials one participant can reasonably complete before fatigue effects increase response variance. 
The cognitive fatigue constraints on the experimental design are important, but otherwise, power calculations for graphics experiments are similar to comparable experiments in other disciplines, in that a pilot study provides the necessary inputs to the sample size calculation for the main experiment.
Some experts have begun to recommend that instead of pilot-study informed power calculations, experimenters use a more general approach based on the intended analysis method [@brysbaertHowManyParticipants2019]. 
As power calculations for various statistical methods can be easily found in any experimental design textbook (such as @easterlingFundamentalsStatisticalExperimental2015), here we provide some basic guidelines based on past experiments.

In psychophysical graphics experiments, such as @luModelingJustNoticeable2022, there are often fewer participants with many more trials per participant (28 participants, 250 trials each); the stimuli in these experiments are often much simpler (e.g. one or two plots instead of 20 in a lineup) and engagement is limited to detection rather than estimation or prediction.
Psychophysics experiments, particularly those designed for analysis using Rasch models, require that each participant assesses a full factorial set of stimuli, whereas analysis with generalized linear mixed models allows for use of incomplete blocks and other strategies that spread the cognitive burden across several participants. 
These experiments have also been traditionally completed in person, which may also explain the large number of trials each participant is asked to complete.

Online graphics experiments often have between 300 and 600 participants [@robinsonPerceptionCognitiveImplications,], but this varies widely with the experimental design, type of stimuli used, and level of participant engagement required; @loyVariationsQQPlots2016 and @clusters had more than 1300 participants,  while @sineillusion and @vanderplasFramedReproducingRevisiting2019 had fewer than 150 participants. 
In-person experiments [e.g. @vanderplasSpatialReasoningData2016] often use fewer participants than online equivalents, but these experiments may require more tasks of each participant, reflecting the increased logistical costs of scheduling participants, supervising the experiment, and entering the data, if tasks are completed on paper. 
In some cases, in-person experiments may combat cognitive fatigue by scheduling the experiment across multiple visits; this leaves experiments vulnerable to participant drop-out effects, but can effectively balance cognitive load and participant recruitment costs in some situations.

# Pilot Testing and Quality Assurance {#sec-pilot-test}

Once the data generating model is set, the protocol is developed, the experiment is designed, and ethics paperwork has been submitted, the next step is another round of testing.
The goal of the initial sequence of testing is to ensure that the experiment is set up properly and that no issues have been overlooked. 
Pilot testing also provides an opportunity to ensure that directions are clear, participants know what they are supposed to be doing, and that the designed study has sufficient power to detect an effect. 
Our studies usually go through 2-3 rounds of preliminary testing, with at least one of those rounds involving any relatives and friends who are less technologically savvy. 
We also purposely include talented individuals who can accidentally crash any testing applications, as a way to harden our data collection software before deployment.

<!-- Pilot study samples do not need to be representative of any particular population, though we would caution against using exclusively visualization researchers or statisticians in your pilot sample because some issues are less likely to show up with knowledgeable participants.  -->

One highly useful (but not strictly essential) component of an experiment that can be set up during the pilot testing stage is a basic analysis script which summarizes all data collected to date visually. 
We have used such scripts in the past to produce automatically updating dashboards or web pages, allowing for real time or near-real time monitoring of data collection efforts. 
This provides an easy way to summarize completion of the experiment so that individuals can receive credit (if using services like MTurk or Prolific), but also allows interested participants to see individual results, which can be a factor when recruiting on social media. 
As data collection online can also happen extremely quickly (300 participants in <2h in our most recent Prolific experiment), this can improve data monitoring, allowing any issues to be spotted and resolved quickly. 
If server load is a potential issue, it may also help to release batches of trials over a longer period of time in order to minimize the chance of having to make participants wait for others to complete the task before the server can handle additional connections^[This is the one major drawback to our preferred solution of self-hosting a Shiny server to handle data collection: the free version of Shiny server is limited to about 15 connections at any given time. Prolific has recently added rate-limiting functions to the experimental control  platform, which makes controlling the number of active jobs much easier.]. 
Batch trial releases can also be used to ensure that participants are recruited across different time zones; in some studies this is beneficial, while in others it may be more important to control the release time to target individuals in e.g. North America rather than Europe.

# Analyzing the Data {#sec-analysis}

One constant with data analysis of these types of experiments is that no matter how carefully the experiment is planned, designed, and executed, there will be surprises. 
This is the dual curse and blessing of studying human perception: the visual system never quite works the way that we expect that it will, which provides endless fodder for science and occasionally complicates the data analysis.

## Generalized Linear Mixed Models
The see-value approach [@niladriroychowdhurySeeValueApp2020;@vanderplasStatisticalSignificanceCalculations2021;@majumderValidationVisualStatistical2013] is extremely useful for single lineups, but when a series of lineups that are part of a designed experiment are used, generalized linear mixed models are a much simpler way to summarize the overall effect of various manipulated factors [@power;@clusters;@emily-diss]. 
This approach also works extremely well for psychometric experiments [@sineillusion], as psychometric models can be easily fit into the framework of a generalized linear mixed model [@juOneModelThat2024] that has higher power than the Rasch models [@andrichRaschModelsMeasurement1988;@luModelingJustNoticeable2022] which were historically recommended for these experiments. 
In addition, similar model structures with different link functions can be used to model accuracy, response time, and confidence, if all three types of information are collected from participants during the experiment.

## Numerical Estimation
There are additional considerations that should be expected when asking participants to estimate numerical quantities. 
Anchoring and rounding cause participant responses to cluster in ways that can bias statistical estimators, requiring methods designed for these types of data [@ushakovStatisticalAnalysisRounded2017;@tourangeau_rips_rasinski_2000;@heitjanIgnorabilityCoarseData1991]. 
An alternative approach is to analyze the data graphically, as shown in @fig-density-rug, which uses a density plot with rug annotations to show individual participant point estimates. 
Rounding effects can clearly be seen in the rug plot, but a kernel density calculated using an appropriately selected bandwidth shows clear visual differences between linear and log scale charts. 
In addition, a smaller second mode can be seen in both linear and log conditions that corresponds to the underlying model value; this suggests that a minority of participants fit a mental regression model to the data and use that model to estimate rather than estimating based off of the closest point. 
When charts are carefully constructed to account for the experimental structure and participant estimation strategies, displaying both individual and aggregate responses, it is possible to demonstrate a measurable difference between conditions without the need for complicated statistical modeling which corrects for rounding and anchoring effects.

While @fig-density-rug shows point estimates, the same approach can be modified to account for experimental methods which generate participant response curves. 
In these cases, rather than adding a rug plot to a density plot showing aggregate estimates, it may be more helpful to create spaghetti plots of individual estimates with a superimposed consensus estimate and relevant annotations showing e.g. anchor points.  
The advantage of this approach is that it can accommodate extremely messy data without requiring the extensive data cleaning, modeling of heuristics like rounding and anchoring, and elimination of nonsensical responses that might be necessary to fit a statistical model. 
While statistical models are undoubtedly beneficial in many situations, we have often found that graphical displays of experimental results are at least as useful for analyzing and presenting the results of graphical experiments.

```{r estimation-data, message=FALSE, warning=FALSE, echo = F, cache = F}
library(tidyverse)
estimation_model_data <- read_csv("https://github.com/earobinson95/EmilyARobinson-UNL-dissertation/raw/main/data/03-estimation/estimation-model-data.csv")
q0_text_summary <- read_csv("https://raw.githubusercontent.com/earobinson95/EmilyARobinson-UNL-dissertation/main/data/03-estimation/q0-text-summary.csv")
estimation_simulated_data <- read_csv("https://github.com/earobinson95/EmilyARobinson-UNL-dissertation/raw/main/data/03-estimation/estimation-simulated-data.csv") %>%
  mutate(x = x - 3000)
estimation_scenario_text <- read_csv("https://github.com/earobinson95/EmilyARobinson-UNL-dissertation/raw/main/data/03-estimation/estimation-scenario-text.csv")
estimation_parameters <- read_csv("https://github.com/earobinson95/EmilyARobinson-UNL-dissertation/raw/main/data/03-estimation/estimation-parameters.csv")
estimation_questions <- read_csv("https://github.com/earobinson95/EmilyARobinson-UNL-dissertation/raw/main/data/03-estimation/estimation-questions.csv")
population_estimates_data <- read_csv("https://github.com/earobinson95/EmilyARobinson-UNL-dissertation/raw/main/data/03-estimation/first-level-population-estimates.csv")

grid_lines_data <- tibble(scale = c(rep("linear", 12), rep("log2", 10)), 
                          grid_lines = c(seq(0,55000, 5000),
                                         2^seq(7,16))
) %>%
  expand_grid(dataset = c("data set1", "data set2"))

qe2_data <- estimation_model_data %>% 
  filter(q_id == "QE2") %>%
  mutate(response = as.numeric(response))
```

```{r density-plot, include = F, echo = F}
densityPlot <- function(data, datasetID, estimate, xlabel = "Estimate", x_limits = c(0,70000), zoom = F, scalesx = T, zoom_limits = c(NA,NA), gridlines = T, rugjitter = 0.25) {

  estPlot <- data %>%
    filter(dataset %in% datasetID) %>%
    ggplot(aes_string(x = estimate, fill = "scale", color = "scale")) +
    geom_density(alpha = 0.5, color = NA) +
    geom_rug(aes(y = -Inf), alpha = 0.6, show.legend = F, position = position_jitter(width = rugjitter, height = 0)) +
    geom_vline(aes(xintercept = true_value, linetype = "a")) +
    geom_vline(aes(xintercept = closest_pt_value, linetype = "b"))

  if (gridlines) {
  estPlot <- estPlot + 
    geom_vline(data = grid_lines_data %>% filter(dataset == "dataset1", grid_lines >= x_limits[1], grid_lines <= x_limits[2]),
               aes(xintercept = grid_lines, color = scale, linetype = "c"))
  }

  estPlot <- estPlot +
    theme_bw() +
    theme(aspect.ratio = 0.5) +
    scale_color_manual("Scale", values = c("steelblue", "orange3")) +
    scale_fill_manual("Scale", values = c("steelblue", "orange3")) +
    scale_linetype_manual("", labels = c("Underlying Model Value", "Closest Point Value", "Grid Line Breaks"), values = c("solid", "dashed", "dotted")) +
    scale_y_continuous("Density", labels = scales::comma)

  if (zoom) {
  estPlot <- estPlot +
    facet_zoom(xlim = zoom_limits)
  }

  if (scalesx) {
    estPlot <- estPlot +
    scale_x_continuous(xlabel, limits = x_limits)
  }

  return(estPlot)

}
```

```{r, qe2-density-plot, message=FALSE, warning=FALSE, echo = F}
#| label: fig-density-rug
#| fig-cap: "Density of participant estimates for the year in which the population reaches 4000. Colors are associated to scale - linear (blue) and log (orange) - and vertical lines indicate the true value based on the underlying model equation (black solid) and closest point value based on the simulated data set (black dashed). A jittered rug plot along the $x$-axis shows where participant estimates were made. The plot shows anchoring occurred to the closest point as shown by an increase in density around the dashed line. Density peaks occurred at whole values indicating rounding errors."
#| fig-width: 6
#| fig-height: 3
#| out-width: "\\columnwidth"

library(patchwork)
# DATA SET 1
qe2_density_1 <- densityPlot(data = qe2_data, datasetID = "dataset1", "response", xlabel = "Estimated Year", x_limits = c(20,35), zoom = F, scalesx = T, gridlines = F, rugjitter = 0.25) +
  ggtitle("Estimation of X given a Y value for Exponential Data") + 
  guides(linetype = guide_legend(title = NULL)) + 
  theme(legend.position = c(1,1), legend.justification = c(1,1), 
        legend.background = element_rect(fill = "white", color = "black"),
        legend.box.just = "right")

qe2_density_1 
```

## Direct Interactions
If participants are making predictions and/or fitting visual statistics, we have had success analyzing these responses by comparing the responses to results from a statistical model to determine how visual statistics differ from the numerical quantities derived mathematically. 
For instance, we calculate the deviation between participant responses and the linear regression in `You Draw It' experiments, then fitted generalized additive mixed models to summarize the results across different experimental conditions to assess how user-drawn predictions deviated from the statistical estimates. 
In other direct interactions, it may be useful to compare participant selections or annotations to closest points on the chart to assess anchoring behavior; for discrete selections, methods discussed in numerical estimation may also be useful.

## Qualitative Responses
In many cases, it is helpful to combine participants' qualitative reasoning with their quantitative responses to designed graphical experiments. 
This approach provides useful context as to what participants use to make their decisions, and can be useful when assessing why unexpected responses occurred. 
Qualitative questions can be as simple as asking for a free response explanation of the quantitative response(s), but it can also be effective to use qualitative questions to prompt participants about problem solving strategies, interesting features in the data, and more.

We have used word clouds to show overall themes in participant explanations of why specific lineup panels were selected [@clusters]; when paired with an appropriate linear model it became clear that participants were fixating on unequal cluster size as a visual cue. 
In cases where participants are provided with additional utilities such as calculators and scratchpads, it can be useful to select responses from individual participants which illustrate the different types of calculations performed, but analyzing this data quantitatively can be difficult, as it may be incomplete or difficult to code systematically.

# Conclusion

Testing features of visualization graphically using online platforms provides an incredibly powerful and efficient way to establish empirical guidelines for statistical graphics and visualization. 
There are nearly endless ways to combine web graphics, user interactions, and data collection to get insight into perception and use of graphics in practical settings. 
We have been continually surprised at the richness of the data collected in these experiments and the ability to combine qualitative and quantitative assessment to support conclusions that are both nuanced and of practical use when deciding how to design and present data using visualizations.

In this paper, we have attempted to contextualize and motivate the logic behind the process we use to design empirical graphics experiments. 
We've discussed model development, experimental design considerations, pilot testing, and data analysis methods that have been honed over many successful and less-than-optimal experiments. 
While no experiment involving humans ever goes exactly to plan, following this process helps to avoid some of the most likely mishaps, ensuring that each new experiment's "bonus" findings have only minimal impacts on the study's overall utility.

While it can be difficult to conduct empirical tests of different visualizations, the results of these experiments support guidelines for graphical design and communication of statistical findings in an accessible and explainable way. 
Many chart design guidelines and recommendations are based on heuristics, but as scientists, we should prefer guidelines which are based on empirical, experimentally derived results over opinions. 
Testing statistical graphics and developing empirically supported guidelines for chart creation promises to support better scientific communication, which is critical for educating the public about topics like climate change, public health, the risk of severe weather and more.

# References